---
title: "Homework 3"
author: "Kelly Chen"
date: "2/9/2022"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**1. Load the prostate cancer data**

```{r}
library(dplyr)
library(glmnet)
library(magrittr)
library(splines)
prostate <- 
  read.table(url(
    'https://web.stanford.edu/~hastie/ElemStatLearn/datasets/prostate.data'))
```

**2. Reproduce correlation matrix**

```{r}
corre<-round(cor(y=prostate[,1:7],x=prostate[,2:8],method="pearson"),3)
corre[upper.tri(corre)]<-""
corre<-as.data.frame(corre)
corre
```



```{r}
prostate
```
**3-4.Treat lcavol as the outcome and Fit Linear Model using lm**

```{r}
# Split Dataset
prostate_train<-prostate %>%
  filter(train==TRUE) %>%
  select(-train)
prostate_test<-prostate %>%
  filter(train==FALSE) %>%
  select(-train)

# Fit Linear Model
fit<-lm(lcavol~.,data=prostate_train)
summary(fit)
```

**5.Use the testing subset to compute the test error (average squared-error loss) using the fitted least-squares regression model**

```{r}
L2_loss<-function(y,yhat){
  (y-yhat)^2
}
error<-function(data,fit,loss=L2_loss){
  mean(loss(data$lcavol,predict(fit,newdata = data)))
}
error(prostate_test,fit)
```

**6.Train a ridge regression model using the glmnet function**

```{r}
#Fit Model
form<- lcavol~lweight+age+lbph+lcp+pgg45+lpsa+svi+gleason
x_inp<-model.matrix(form,data=prostate_train)
y_out<-prostate_train$lcavol
fit<-glmnet(x=x_inp,y=y_out,lambda=seq(0.5,0,-0.05),alpha=0)
fit$beta


```

```{r}
#Train Error
error_g<-function(dat,fit,lam,form,loss=L2_loss){
  x_inp<-model.matrix(form,data=dat)
  y_out<-dat$lcavol
  y_hat<-predict(fit,newx=x_inp,s=lam)
  mean(loss(y_out,y_hat))
}

# Guess and Check


error_g(prostate_train,fit,lam = 0.18,form=form)

error_g(prostate_test,fit,lam =0.18,form=form)


```

**Step 7. Create figure of test error and training error**

```{r}
# Compute training and test errors
error_train<-sapply(fit$lambda,function(lamb)
  error_g(prostate_train,fit,lamb,form))
error_test<-sapply(fit$lambda,function(lamb)
  error_g(prostate_test,fit,lamb,form))

#Plot
plot(x=range(fit$lambda),
     y=range(c(error_test,error_train)),
     xlim=rev(range(fit$lambda)),
     type='n',
     xlab=expression(lambda),
     ylab='train/test error')
points(fit$lambda,error_train,pch=19,type='b',col='blue')
points(fit$lambda,error_test,pch=19,type='b',col='red')
legend('topright',c('train','test'),lty=1,pch=19,col=c('blue','red'),bty='n')
#colnames(fit$beta)
```






**Step 8. Create figure for path diagram of the ridge regression analysis**

```{r}
plot(x=range(fit$lambda),
     y=range(as.matrix(fit$beta)),
     type='n',
     xlab=expression(lambda),
     ylab='coefficients')
for(i in 1:nrow(fit$beta)){
  points(x=fit$lambda,y=fit$beta[i,],pch=10,col='red')
  lines(x=fit$lambda,y=fit$beta[i,],col='green')
}
text(x=0,y=fit$beta[,ncol(fit$beta)],
     labels=rownames(fit$beta),
     xpd=NA,pos=4,srt=45)
abline(h=0,lty=3,lwd=1)
```

