---
title: "Homework-5"
author: "Kelly Chen"
date: "3/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Original Code

```{r}
library('MASS') ## for 'mcycle'
library('manipulate') ## for 'manipulate'
library(caret)

y <- mcycle$accel
x <- matrix(mcycle$times, length(mcycle$times), 1)

plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")

## Epanechnikov kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## lambda - bandwidth (neighborhood size)
kernel_epanechnikov <- function(x, x0, lambda=1) {
  d <- function(t)
    ifelse(t <= 1, 3/4*(1-t^2), 0)
  z <- t(t(x) - x0)
  d(sqrt(rowSums(z*z))/lambda)
}

## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}

## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}

## Helper function to view kernel (smoother) matrix
matrix_image <- function(x) {
  rot <- function(x) t(apply(x, 2, rev))
  cls <- rev(gray.colors(20, end=1))
  image(rot(x), col=cls, axes=FALSE)
  xlb <- pretty(1:ncol(x))
  xat <- (xlb-0.5)/ncol(x)
  ylb <- pretty(1:nrow(x))
  yat <- (ylb-0.5)/nrow(x)
  axis(3, at=xat, labels=xlb)
  axis(2, at=yat, labels=ylb)
  mtext('Rows', 2, 3)
  mtext('Columns', 3, 3)
}

## Compute effective df using NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## kern  - kernel function to use
## ... - arguments to pass to kernel function
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}

## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2

## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

## AIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

## BIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d


## make predictions using NW method at training inputs
y_hat <- nadaraya_watson(y, x, x,
  kernel_epanechnikov, lambda=5)

## view kernel (smoother) matrix
matrix_image(attr(y_hat, 'k'))

## compute effective degrees of freedom
edf <- effective_df(y, x, kernel_epanechnikov, lambda=5)
aic(y, y_hat, edf)
bic(y, y_hat, edf)

## create a grid of inputs 
x_plot <- matrix(seq(min(x),max(x),length.out=100),100,1)

## make predictions using NW method at each of grid points
y_hat_plot <- nadaraya_watson(y, x, x_plot,
  kernel_epanechnikov, lambda=1)

## plot predictions
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot, y_hat_plot, col="#882255", lwd=2) 

## how does k affect shape of predictor and eff. df using k-nn kernel ?
# manipulate({
#   ## make predictions using NW method at training inputs
#   y_hat <- nadaraya_watson(y, x, x,
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   edf <- effective_df(y, x, 
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   aic_ <- aic(y, y_hat, edf)
#   bic_ <- bic(y, y_hat, edf)
#   y_hat_plot <- nadaraya_watson(y, x, x_plot,
#     kern=kernel_k_nearest_neighbors, k=k_slider)
#   plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
#   legend('topright', legend = c(
#     paste0('eff. df = ', round(edf,1)),
#     paste0('aic = ', round(aic_, 1)),
#     paste0('bic = ', round(bic_, 1))),
#     bty='n')
#   lines(x_plot, y_hat_plot, col="#882255", lwd=2) 
# }, k_slider=slider(1, 10, initial=3, step=1))


```


# Question 1: Split the data into 75% and 25%

```{r}
library(dplyr)
set.seed(123)
train<-sample_frac(mcycle,size=0.75)
test<-anti_join(mcycle,train)
train_x<-matrix(train$times)
test_x<-matrix(test$times)
train_y<-train$accel
test_y<-test$accel

class(train_x)

```

# Question 2: Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values.

```{r}

## plot predictions
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot, y_hat_plot, col="#882255", lwd=2)


# Use manipulate to check parameters
#manipulate({
 # y_hat_plot<- nadaraya_watson(y, x, x_plot,
    #kern=kernel_k_nearest_neighbors, k=k_slider)
 # plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
  #lines(x_plot, y_hat_plot, col="#882255", lwd=2) 
#}, k_slider=slider(1, 100, initial=1, step=1))

# Set k=11
y_hat_12<- nadaraya_watson(y, x, x_plot,
    kern=kernel_k_nearest_neighbors, k=11)

plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot, y_hat_12, col="#882255", lwd=2)
```

I used manipulate to find the best k that I think would fit the model when k is 11. 


# Question 3: With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.

```{r}
# Train Error
y_hat<-nadaraya_watson(train_y,train_x,train_x,kern=kernel_k_nearest_neighbors,k=1)
d<-effective_df(train_y, train_x, kern=kernel_k_nearest_neighbors,k=1)
aic_test<-aic(train_y,y_hat,d=d)
bic_test<-bic(train_y,y_hat,d=d)
train_err<-error(train_y,y_hat)

#Validation Error
y_hat_test<-nadaraya_watson(test_y,test_x,test_x,kern=kernel_k_nearest_neighbors,k=1)
validation_err<-error(test_y,y_hat_test)

# Plot
y_hat_plot<-nadaraya_watson(train_y, train_x, x_plot,
    kern=kernel_k_nearest_neighbors, k=1)


plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
legend('bottomright', legend = c(
  paste0('eff. df = ', round(d,1)),
  paste0('train error = ', round(train_err,1)),
  paste0('validation error= ', round(validation_err,1)),
  paste0('aic = ', round(aic_test, 1)),
  paste0('bic = ', round(bic_test, 1))),
  bty='n')
lines(x_plot, y_hat_plot, col="#882255", lwd=2) 


# Manipulate
#manipulate({
  #y_hat<-nadaraya_watson(train_y,train_x,train_x,kern=kernel_k_nearest_neighbors,k=k_slider)
  #d<-effective_df(train_y, train_x, kern=kernel_k_nearest_neighbors, k=k_slider)
  #aic_test<-aic(train_y,y_hat,d)
  #bic_test<-bic(train_y,y_hat,d)
  #train_err<-error(train_y,y_hat)
  #y_hat_test<-nadaraya_watson(test_y,test_x,test_x,kern=kernel_k_nearest_neighbors,k=k_slider)
  #validation_err<-error(test_y,y_hat_test)
 # y_hat_plot<-nadaraya_watson(train_y, train_x, x_plot,kern=kernel_k_nearest_neighbors, k=k_slider)
  #plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
 #legend('bottomright', legend = c(
    #paste0('eff. df = ', round(d,1)),
    #paste0('train error = ', round(train_err,1)),
   # paste0('validation error= ', round(validation_err,1)),
   # paste0('aic = ', round(aic_test, 1)),
   # paste0('bic = ', round(bic_test, 1))),
   # bty='n')
  #lines(x_plot, y_hat_plot, col="#882255", lwd=2) 
#},k_slider=slider(1,10,initial=1,step=1))
```

After using manipulation, I found out the validation error is lowest when it is at k=1.

# Question 4: For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.
```{r}
set.seed(123)
fold_y <-createFolds(mcycle$accel,k=5)
print(fold_y)
sapply(fold_y, length)

cvknnreg <- function(kNN = 10, flds=fold_y) {
  cverr <- rep(NA, length(flds))
  for(tst_idx in 1:length(flds)) { ## for each fold
    
    ## get training and testing data
    mcycle_trn <- mcycle[-flds[[tst_idx]],]
    mcycle_tst <- mcycle[ flds[[tst_idx]],]
    
    ## fit kNN model to training data
    knn_fit <- knnreg(accel~times,
                      k=kNN, data=mcycle_trn)
    
    ## compute test error on testing data
    pre_tst <- predict(knn_fit, mcycle_tst)
    cverr[tst_idx] <- mean((mcycle_tst$accel - pre_tst)^2)
  }
  return(cverr)
}


## Compute 5-fold CV for kNN = 1:20
cverrs <- sapply(1:20, cvknnreg)
print(cverrs) ## rows are k-folds (1:5), cols are kNN (1:20)


cverrs_mean <- apply(cverrs, 2, mean)
cverrs_sd   <- apply(cverrs, 2, sd)
```


# Question 5: Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).

```{r}
plot(x=1:20, y=cverrs_mean, 
     ylim=range(cverrs),
     xlab="'k' in kNN", ylab="CV Estimate of Test Error")
segments(x0=1:20, x1=1:20,
         y0=cverrs_mean-cverrs_sd,
         y1=cverrs_mean+cverrs_sd)
best_idx <- which.min(cverrs_mean)
points(x=best_idx, y=cverrs_mean[best_idx], pch=20)
abline(h=cverrs_mean[best_idx] + cverrs_sd[best_idx], lty=3)
```


# Question 6: Interpret the resulting figures and select a suitable value for the tuning parameter.

The resulting figure shows that the minimum test error happens when k=10, and also when k is larger than 10 the test error gradually increases, so I would choose k=10 as the tuning parameter.

